Wikipedia Text Mining and Entropy Analysis

### 2018-09-26 wiki_entropy.Rmd
changes to wiki_entropy
wiki_entropy slope analysis not function, cannot fix error at present

def get_sen_df(text):
	sd = get_sen_dict(text)
	s = pd.Series(sd, name = "length")
	s.index.name = "sentence"
	df = s.reset_index() #gives a dataframe with columns "sentence" and "length" containing the sentence 
					 	  #and the setence length respectively
	return df

text = extract_texts(50)
ro_df = get_sen_df(text)
```

```{r analyze_entropies}
get_entropies <- function(sen_length, df) {

@@

entropies <- map(6:12, ~get_entropies(.x, py$ro_df)) %>%
  bind_rows()

ggplot(entropies, aes(x = word_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F)
```

>>>>

def get_sen_df(text):
	sd = get_sen_dict(text)
	s = pd.Series(sd, name = "length")
	del sd
	s.index.name = "gloss"
	df = s.reset_index() #gives a dataframe with columns "sentence" and "length" containing the sentence 
					 	  #and the setence length respectively
	df["speaker_role"] = "not child"
	return df

text = extract_texts(20)
xmf_df = get_sen_df(text)
```

```{r slope_code, include=FALSE}
extract_stem_tokens <- function(role, min_length, utterances) {

  if (role == "child") {
    lengths <- utterances %>%
      filter(speaker_role == "Target_Child") %>%
      mutate(length = str_count(stem, " ") + 1)

} else {
    lengths <- utterances %>%
      filter(speaker_role != "Target_Child") %>%
      mutate(length = str_count(stem, " ") + 1)
}

  tokens <- lengths %>% 
    filter(length >= min_length) %>%
    mutate(utterance_id = 1:n()) %>%
    unnest_tokens(word, stem) %>%
    group_by(utterance_id) %>%
    mutate(word_order = 1:n()) %>%
    group_by(length, word_order, word) %>%
    summarise(n = n()) 
}

extract_gloss_tokens <- function(role, min_length, utterances) {

  if (role == "child") {
    lengths <- utterances %>%
      filter(speaker_role == "Target_Child") %>%
      mutate(length = str_count(gloss, " ") + 1)

} else {
    lengths <- utterances %>%
      filter(speaker_role != "Target_Child") %>%
      mutate(length = str_count(gloss, " ") + 1)
}

  tokens <- lengths %>% 
    filter(length >= min_length) %>%
    mutate(utterance_id = 1:n()) %>%
    unnest_tokens(word, gloss) %>%
    group_by(utterance_id) %>%
    mutate(word_order = 1:n()) %>%
    group_by(length, word_order, word) %>%
    filter(word_order <= length) %>% 
    summarise(n = n()) 
}

start_slope <- function(df) {

 df %>%
    filter(word_order %in% c(1,2)) %>%
    sample_frac(1, replace = T) %>%
    group_by(length, word_order) %>%
    summarise(entropy = entropy(n, unit = "log2")) %>%
    spread(word_order, entropy) %>%
    ungroup() %>%
    mutate(diff = `2` - `1`) %>%
    summarise(diff = mean(diff)) %>%
    pull(diff)

}

second_slope <- function(df) {

 df %>%
    ungroup() %>%
    mutate(word_order %in% c(2, 3)) %>%
    filter(word_order > 0) %>%
    group_by(length, word_order) %>%
    sample_frac(1, replace = T) %>%
    group_by(length, word_order) %>%
    summarise(entropy = entropy(n, unit = "log2")) %>%
    spread(word_order, entropy) %>%
    ungroup() %>%
    mutate(diff = `3` - `2`) %>%
    summarise(diff = mean(diff)) %>%
    pull(diff)

}

#note this measurement only makes sense if utterance length is at least 6
mid_slope <- function(df) {

 df %>%
    ungroup() %>%
    mutate(word_order = if_else(word_order == length-2, 2,
                                if_else(word_order == 3, 1,
                                        0))) %>%
    filter(word_order > 0) %>%
    group_by(length, word_order) %>%
    sample_frac(1, replace = T) %>%
    group_by(length, word_order) %>%
    summarise(entropy = entropy(n, unit = "log2")) %>%
    spread(word_order, entropy) %>%
    ungroup() %>%
    mutate(diff = `2` - `1`) %>%
    summarise(diff = mean(diff)) %>%
    pull(diff)

}

penult_slope <- function(df) {

 df %>%
    ungroup() %>%
    mutate(word_order = if_else(word_order == length-1, 2,
                                if_else(word_order == length -2, 1,
                                        0))) %>%
    filter(word_order > 0) %>%
    group_by(length, word_order) %>%
    sample_frac(1, replace = T) %>%
    group_by(length, word_order) %>%
    summarise(entropy = entropy(n, unit = "log2")) %>%
    spread(word_order, entropy) %>%
    ungroup() %>%
    mutate(diff = `2` - `1`) %>%
    summarise(diff = mean(diff)) %>%
    pull(diff)

}

end_slope <- function(df) {

 df %>%
    ungroup() %>%
    mutate(word_order = if_else(word_order == length, 2,
                                if_else(word_order == length -1 , 1,
                                        0))) %>%
    filter(word_order > 0) %>%
    group_by(length, word_order) %>%
    sample_frac(1, replace = T) %>%
    group_by(length, word_order) %>%
    summarise(entropy = entropy(n, unit = "log2")) %>%
    spread(word_order, entropy) %>%
    ungroup() %>%
    mutate(diff = `2` - `1`) %>%
    summarise(diff = mean(diff)) %>%
    pull(diff)

}

get_slopes <- function(role, min_length, utterances, stemgloss) {

  if (stemgloss == "stem") {
    tokens <- extract_stem_tokens(role, min_length, utterances)  
  } else {
    tokens <- extract_gloss_tokens(role, min_length, utterances)
  }

  start_slopes <- replicate(1000, start_slope(tokens)) %>% 
    as_data_frame() %>%
    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
              ci_lower = quantile(value, .025))

  second_slopes <- replicate(1000, second_slope(tokens)) %>% 
    as_data_frame() %>%
    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
              ci_lower = quantile(value, .025))

  mid_slopes <- replicate(1000, mid_slope(tokens)) %>% 
    as_data_frame() %>%
    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
              ci_lower = quantile(value, .025))

  penult_slopes <- replicate(1000, penult_slope(tokens)) %>% 
    as_data_frame() %>%
    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
              ci_lower = quantile(value, .025))

  end_slopes <- replicate(1000, end_slope(tokens)) %>% 
    as_data_frame() %>%
    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
              ci_lower = quantile(value, .025))

  rbind(start_slopes, second_slopes, mid_slopes, penult_slopes, end_slopes)
}

```

```{r slope_analysis}
#get_slopes("not child", 6, py$xmf_df, "gloss")
gtx <- extract_gloss_tokens("not child", 6, py$xmf_df)
start_slopes <- replicate(1000, start_slope(gtx)) %>% 
    as_data_frame() %>%
    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
              ci_lower = quantile(value, .025))
print(start_slopes)
```

```{r analyze_entropies, include=FALSE}
get_entropies <- function(sen_length, df) {

@@

#entropies <- map(6:12, ~get_entropies(.x, py$xmf_df)) %>%
#  bind_rows()

#ggplot(entropies, aes(x = word_order, y = empirical_entropy,
#                      ymin = ci_lower, ymax = ci_upper)) +
#  facet_wrap(~ length) + 
#  geom_pointrange() +
#  geom_smooth(se = F)
```
